Modular Synthesis of Verified Verifiers of Computation with STV Algorithms

### 2019-03-12 paper/paper.tex
the tex source file updated

\maketitle
\begin{abstract}
Single transferrable vote (STV) is a family of preferential voting systems, different instances of which are used in binding elections throughout the world. We give a formal specification of this family, from which we derive fully verified tools that verify the computation for various instances of STV vote counting. These tools validate the probably correct execution of a run of a vote counting algorithm, based on a transcript of the count.

Our framework distils the similarities and differences of various instances of STV and gives a uniform and modular way of synthesising verifiers for its various instances, and provides the flexibility and ease for adapting and extending it to a variety of STV schemes. We minimise the trusted base in correctness of the tools produced by using the HOL4 and CakeML as the technical basis. We first formally specify and verify the tools in HOL4 and then obtain the machine executable versions for the tools by relying on the verified proof translator and the compiler of the CakeML. Moreover, proofs that we establish in HOL4 and CakeML are almost completely automated so that new verified instances of STV can be created with no (or minimal) extra proof. Finally, our experimental results with executable code demonstrate feasibility of deploying the framework for verifying real size elections having an STV counting algorithm.
\end{abstract}

@@

The goal of this paper is to advance formal methods in elections by demonstrating how STV vote counting can be verified in a modular, transparent, and practical feasible way over a minimal computational trust base:  we show how to synthesise efficient and provably correct verifies for a large class of STV schemes in a modular and almost fully automated.

@@

We achieve this by first giving a formal definition STV (called \emph{generic STV}) 
that  encompasses the entire family of STV vote counting
schemes. This definition necessarily under-specifies any particular STV vote counting scheme, but fixes both basic data structures and mode of computation common to all instances of STV. Trustworthiness is achieved by
not verifying the \emph{software} that computes the result, but instead by verifying \emph{evidence} that substantiates the correctness of the count. Evidence records and corresponds to the individual steps taken in the computation of winners 
%(That is, our approach is an instance of verifiable computation \footnote{DP: reference}). 
%The evidence directly corresponds to the individual steps of the count
(and could so also be verified, or spot-checked manually), giving transparency. We minimize the computational trust base by verifying our executables down to the machine level using a combination of HOL \cite{hol} and CakeML \cite{cake}.

@@

To begin 
%with addressing the problem, 
assume $\mathcal{P}$ is an implementation of an algorithm $\mathcal{A}$. We reasonably demand each execution of $\mathcal{P}$ on a given input $x$ consisting of ballots recorded in the election to output the winners $y$ and evidence $\omega$ as claim  for correctness of the execution.  
 Having evidence available for each execution of $\mathcal{P}$ facilitates checking correctness of the computation carried out \emph{independently} of the (source code of) $\mathcal{P}$. Therefore the original problem boils down to how one can verify the evidence of any \emph{instance of computation} with any algorithm $\mathcal{A}$.

The evidence as such must be enough informative to provide transparency of tallying and also allow  voters, or at least a large pool of scrutineers, to verify for themselves that the tally is authentically processed as per instructions of the counting scheme  so that winners truly reflect will of the voters.    
Therefore the questions becomes what information is enough  
for establishing transparency and verifiability of tallying  and what data structure for recording the information in evidence we should choose.

@@

 \item\textbf{elected candidates} who are the already elected ones whose surplus votes must be transferred,

@@

In a secondary examination of the STV family, we also identify a general algorithmic mechanism for performing computation on evidence, which realises the \emph{commonalities} among the family members. This generic method comprises seven general actions~(Figure~\ref{GenSteps}) corresponding to actions that tally officers take when counting votes.
\begin{figure} 
 \begin{itemize}
\item\textbf{count} for counting the uncounted ballots,
\item\textbf{elect} for electing all or some of the electable candidates, 
\item\textbf{transfer-elected} for distributing surplus votes of elected, %candidates, 
\item\textbf{eliminate} to exclude one or some candidates, 
\item\textbf{transfer-removed} to distribute votes of   the eliminated, %candidates, 
\item\textbf{elected-win} for terminating tallying whenever the vacancies are filled or there is no continuing candidate left, 
\item\textbf{hopeful-win} for terminating tallying whenever  the number of already elected and continuing candidates collectively does not exceed the initial vacancies. 
\end{itemize}
\caption{Generic STV Counting Steps}
\label{GenSteps}
\end{figure}
Each action is constituted of  pre-conditions on when and how the action comes into effect, and post-conditions dictating what effect applying the action brings about by describing the new state to which computation proceeds. In fact, these conditions are the formal counterparts of the legal clauses in the textual specification of the counting scheme which tell tally officers how to proceed with tallying, and they are common between all flavours of STV.

For example a constant pre-condition for applying elimination  asserts that there must be no uncounted ballot left and another one declares that no candidate should be electable at the current state of counting. A post-condition of elimination is that once a candidate is excluded from counting, the list of continuing candidates is updated accordingly so that the excluded candidate no longer received votes. Conjunctions of such declarations comprise what an action is and what an application of that action means.

@@

\begin{definition}[The generic STV machine]\label{STVMachine}
Let $\mathcal{S}$  and $\mathcal{T}$
be respectively the set of machine states and transition labels.  Also assume for $t\in\mathcal{T}$, $S_{t}$ $=$ $\bigwedge_{\atop{i\leq j_{t}}} \psi_{i}$ where  $\psi_{i}$ is the formal specification (in higher-order logic) of a pre- or post-condition of $t$. Then \emph{generic STV} is the triple $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$.  
\end{definition}
There are nonetheless variations in STV algorithms  which we  recognise  by separate \emph{instantiations} into the machine.  
\subsection{Instantiations of the Machine with Specific STV schemes}
Differences exist between counting algorithms of STV schemes.  
 For example in the STV used for electing the upper house representatives of the Victoria state of Australia (Victoria STV), distributing votes of an eliminated candidate occurs step by step in order of the magnitude of the fractional values that those votes carry.  Consequently, a pre-condition for applying the action is to first assure that upon each instance of applying transfer-removed action (a) exactly those votes which have the same fractional value are transferred and (b) the transfer happens according to the magnitude order. In contrast, for example, the STV used in the lower house elections of Tasmania state of Australia (Tasmania STV) does not have such pre-condition simply because votes of an eliminated candidate are transferred in a single application of transfer-removed. There are also variations in post-conditions. For example transfer-elect of the ACT STV used in the lower house elections of the ACT state of Australia requires distributing only the last parcel of surplus votes received by an elected candidate. But Victoria STV specifies transferring all surplus votes of an elected candidate.

We formally realise variations of STV algorithms by instantiation~(Definition\ref{STVInst}) of the generic machine. Recall that the semantics of the machine is formed by conjunctions of formally specified universally appearing pre and post-conditions in STV algorithms. An instantiation of the machine with an STV algorithm $\mathcal{A}$ happens by enriching the generic semantics by adding formal specifications of the legal clauses that are specific to $\mathcal{A}$. We hence come to define an operational semantics functioning in accordance with instructions of $\mathcal{A}$. 
\begin{definition}[Instantiation of the Machine]\label{STVInst}
Assume the generic machine $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$ as in Definition~\ref{STVMachine}. An instantiation $\hat{\mathcal{A}}$ for an STV algorithm $\mathcal{A}$ into the generic STV machine $M$ is a triple $\langle \mathcal{S}, \mathcal{T}, (S'_t)_{t \in \mathcal{T}} \rangle$, where for each $t\in\mathcal{T}$, $S'_{t}$ $=$ $S_{t}\wedge\bigwedge_{\atop{i\leq j'_{t}}} \phi_{i}$. Each $\phi_{i}$ is the formal specification of a pre- or post-condition specific to $\mathcal{A}$.   
\end{definition}

The operational semantics of each instantiation allows  to perform operations such as validation on evidence. Definition~\ref{verifier} lays down what verifying an evidence amount to. Informally speaking, to check if  evidence is valid one simply needs to inspect if transitions  between each two consecutive states of computation appearing in the evidence occur by a legitimate application of a counting action of the scheme used.

\begin{definition}[verifier]\label{verifier}
Assume $\hat{\mathcal{A}}= \langle \mathcal{S}, \mathcal{T}, (S'_t)_{t \in \mathcal{T}} \rangle$ is an instantiation of the STV algorithm $\mathcal{A}$. A verifier for instances of computation with $\mathcal{A}$ is a function $\hat{\mathcal{V}}$ mapping from $\mathbb{Q}\times\mathbb{N}\times 2^{C}\times$ \textsf{List}$(\mathcal{S})$ to $\{0,1\}$ such that for any evidence $\hat{\omega} = (Q,s,C,\Omega)$ where $\Omega=\langle\Omega_{1},\dots,\Omega_{n}\rangle$, $\hat{\mathcal{V}}(\hat{\omega}) = 1$ if and only if the following holds:\\
$\forall i\in\{1,\dots,n-1\}.$ 
%\hspace*{0.2cm}$\forall \Omega_{i-1}$ $\Omega_{i}\in\mathcal{S}.$
%$\exists \Omega_{1}$ $\Omega_{2}\in$ \textsf{List} $(\mathcal{S}).$ $\Omega = \Omega_{1} + [\Omega_{i-1};\Omega_{i}] + \Omega_{2}$  
$\exists t\in\mathcal{T}.$  $(\Omega_{i-1},\Omega_{i})\in t$ $\wedge~\vdash\mathcal{S}'_{t}[\Omega_{i-1},\Omega_{i}]$\\ 
where $\mathbb{Q}$ and $\mathbb{N}$ respectively represent the set of rational and natural numbers, $2^{C}$ is the set of all subsets of $C$, and \textsf{List}$(S)$ is the set of all possible lists of machine states. $\Omega_{i-1}$ is a pre-state and $\Omega_{i}$ is a post-state visited in the execution. The last line means that pre- and post-conditions  of applying $t$ to move from $\Omega_{i-1}$ to $\Omega_{i}$  are logically true. 
\end{definition} 
We shall next provide an example to concretely illustrate a piece of evidence and exemplify how STV schemes work. 
 \subsection{An Example of a Concrete Piece of Evidence}
\label{subsec:InstEv}
Figure~\ref{EvInst} depicts evidence of a small election whose counting algorithm is the ACT STV 
with three candidates $a$, $b$ and $c$, and initially recorded ballots containing
$b_1=([a,c],1/1)$, $b_2=([a,b,c],1/1)$, $b_3=([a,c,b],1/1)$,
$b_4=([b,a],1/1)$, $b_5=([c,b,a],1/1)$.  The first component of each ballot is a preference-ordered list of candidates, and the second is the fractional transfer value (initially set to $1/1$). The evidence consists of a \emph{header} that specifies the quota (computed according to the ACT STV instructions), the number of vacancies, and the list of competing candidates. 
The fourth line is the election result, and the remainder of the evidence are non-final states visited to compute this
outcome. Each non-final state shows uncounted ballots, tallies of candidates, piles of votes attracted, backlog of the elected, backlog of the eliminated, list of elected candidates, and list of continuing candidates.

First preferences for each candidate are computed, and ballots counted in favour of particular candidates are placed onto that candidate’s pile. Here, $a$ is the first preference on $b_{1}$, $b_{2}$, and $b_{3}$, and $b$ receives $b_{4}$, and $c$ receives $b_{5}$ . Tallies are updated so that tally of $a$ becomes 3, and $b$
and $c$ each reach $1/1$. As $a$ exceeds the quota, he is elected and the fractional value of his surplus is updated according to the ACT scheme. Then $a$'s surplus is transferred to continuing candidates $b$ and $c$ and votes are counted according to next preferences shown on $a$'s surplus votes. Candidate $b$ and $c$ each attracts one vote. However  neither of them reaches to the quota. Therefore the weakest one $b$ is eliminated. As the number of elected and continuing candidates does not exceed the initial vacancies, the election terminates by declaring $a$ and $c$ winners.
\begin{small} 
\begin{figure}[b]
\begin{tabular}{c@{\hspace{2cm}}c} 
\AxiomC{\tiny 8/3} \noLine

@@

\UnaryInfC{\tiny [$b_4$,([a,b,c],1/9)]]; a\{3/1\} b\{10/9\} c\{11/9\}; a\{[]\} b\{[]\}
c\{[[$b_{5}$,([c],1/9),([c,b],1/9),([c],1/9)]]\}; []; [a]; [c]}

@@

   \node[state,
  below of= STV1,
  xshift = - 1cm,
  node distance=1.2cm,
  anchor=center] (T) 
 {%
 \begin{tabular}{l}
  \textbf{Tran}
 \end{tabular}
 };

@@

We also have another motivation for developing the auxiliary module through this three-part process of specification, implementation and then verification. Familiarity, but no expertise, with our framework and its general purpose is needed for extending it to synthesise an executable verifier for one's desired STV scheme. On the one hand, an average user may lack enough skills in grasping functional programming style which we rely on for implementing verifiers in HOL4. However, the same user most probably has had exposure to formulations of mathematical properties  in first-order logic syntax. Including purely descriptive logical assertions as our specifications, proven equivalent to their implementations, facilitates the users with means to understand the functionality of components, modules and the system as a whole simply by inspecting the specifications instead of implementations.
\subsection{Data Structure of Machine States}
\label{MachineData}
 We choose the data structure  given in Definition~\ref{judgement} for implementing the abstract data type underlying the generic STV.  We have four reasons for this choice of data structure. 
 \begin{itemize}
\item As the abstract syntactic representation of evidence closely represents concrete evidences, designing the parser and inspecting  its  correctness is less challenging.  
 \item HOL4 has well developed libraries comprised of verified assertions of operations on list structure. By structuring the data on lists, we facilitate us and users  with exploitation of  already verified tools to formalise the framework and avoid inventing the unnecessary from the scratch. 
 \item HOL4 also has well-developed tactics for discharging proof obligations on assertions involving list structure and operations on it. We therefore come to provide us with means for ease of verification of the formalised assertions.
 \item Understanding the data structure used in the  framework implementation is critical for ease of  its usability and extensibility by third parties.  The type $\mathit{judgement}$ closely models concrete evidences such as the one in figure~\ref{EvInst}. Therefore one can sensibly perceive the  abstraction step taken for modelling  concrete data which in turn enhances understandability of the framework and its mechanism. 
   \end{itemize}
\begin{definition}\label{judgement}
We formalise machine states as an inductive type  $\mathit{judgement}$ whose constructors are  $\mathit{NonFinal}$ and $\mathit{Final}$; 
\begin{holthmenv}
\HOLTyOp{judgement}\;=\\
\;\;\;\;\HOLConst{NonFinal}\;(\HOLTyOp{ballots}\;\HOLTokenProd{}\\
\;\;\;\;\;\HOLTyOp{tallies}\;\HOLTokenProd{}\\
\;\;\;\;\;\HOLTyOp{piles}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list})\\
\;\;\HOLTokenBar{}\;\HOLConst{Final}\;(\HOLTyOp{cand}\;\HOLTyOp{list})
\end{holthmenv}

 A $\mathit{Final~(w)}$ judgement declares $w$ as winners of the election. A  $\mathit{NonFinal~(ba,t,p,bl_{1},bl_{2},e,h)}$ judgement  consists of uncounted ballots $ba$, tally $t$, pile $p$, $bl_{1}$ the backlog of elected candidates, $bl_{2}$ the backlog of eliminated candidates, and $e$ and $h$ for the list of elected and continuing~(hopeful) candidates. The types  $\mathit{ballots}$, $\mathit{tallies}$ and $\mathit{piles}$ are respectively abbreviations for  $\mathit{((cand ~list)\times rat) list}$,   $\mathit{(cand\times rat)~list}$ and $\mathit{(cand\times(ballots~list)~list) ~list}$ where $\mathit{rat}$ is the HOL4 type of fractional numbers.
\end{definition}
 Recall that objects whose type is $\mathit{piles}$ serve as containers for recording the ballots allocated to each candidate. Our choice for type of piles allows us to store the ballots received by each candidate in chunks of lists rather than one single list containing all of them.  We have two reasons for designing the type of piles as it stands instead of $\mathit{(cand\times(ballots~list))~list}$;  
\paragraph*{(a)} Some STV schemes such as the lower house ACT and Tasmania STV, employ a nation called ''last parcel''. In short the last parcel of a candidate is the collection of those ballots received by the candidate at the last round of application of the count transition which made the tally of the candidate reach or exceed the quota and therefore be elected.  Only the last parcel, instead of all, of an elected candidate is then distributed the ballots received by the candidate. Also  they update the fractional transfer value, that each ballot in the last parcel  awaiting transfer carries, based on the length of the last parcel. To  accommodate this notion and its effect, we need identifying the ballots constituting the last parcel of an elected. We hence formalise the pile to record, for each  continuing candidate, a list consisting of lists of ballots each received upon  chronological applications of count transition.  Once a candidate is elected their pile may be manipulated differently from the continuing candidates. 
\paragraph*{(b)} We noted earlier that the STV used for the upper house elections in the Victoria state transfers votes of an eliminated candidate chunk by chunk in several applications of transfer-removed, rather than in a single action, in order of the magnitude of the fractional value that the chunks carry. We therefore need to rearrange the pile of ballots of an eliminated candidate into lists based on equality of the fractional value of ballots and then distribute them in the proper order as the scheme requires. This means that the type of piles has to be as it is defined above.   
Implementing piles in this way enables us to tailor the semantics of the transfer and elect transitions and their instantiations in such a way to modularly formalise several STV schemes which use the last parcel or stepwise distribution of votes form eliminated candidates or the combination of both.

@@

The predicate Valid\textunderscore{}Init\textunderscore{}CandList which realises item (6) asserts that each candidate has a pile and a tally. We also declare that elements in the first component of the tally $t$ and pile $p$ are distinct. Therefore we come to satisfy item (1).

>>>>

\maketitle
\begin{abstract}
Single transferable vote (STV) is a family of preferential voting systems, different instances of which are used in binding elections throughout the world. We give a formal specification of this family, from which we derive fully verified tools that verify the computation for various instances of STV vote counting. These tools validate 
%the probably correct execution of a run of a 
 an execution of an STV vote counting algorithm, based on a transcript of the count.

Our framework distils the similarities and differences of various instances of STV and gives a uniform and modular way of synthesising verifiers for its various instances, and provides the flexibility and ease for adapting and extending it to a variety of STV schemes. We minimise the trusted base in correctness of the tools produced by using the HOL4 and CakeML as the technical basis. We first formally specify and verify the tools in HOL4 and then obtain the machine executable versions for the tools by relying on the verified proof translator and the compiler of CakeML. Moreover, proofs that we establish in HOL4 and CakeML are almost completely automated so that new verified instances of STV can be created with no (or minimal) extra proof. Finally, our experimental results with executable code demonstrate feasibility of deploying the framework for verifying real size elections having an STV counting algorithm.
\end{abstract}

@@

The goal of this paper is to advance formal methods in elections by demonstrating how STV vote counting can be verified in a modular, transparent, and practical feasible way over a minimal computational trust base:  we show how to synthesise efficient and provably correct verifiers for a large class of STV schemes in a modular and almost fully automated.

@@

We achieve this by first giving a formal definition of STV (called \emph{generic STV}) 
that  encompasses the entire family of STV vote counting
schemes. This definition necessarily under-specifies any particular STV vote counting scheme, but fixes both basic data structures and the mode of computation common to all instances of STV. Trustworthiness is achieved by
not verifying the \emph{software} that computes the result, but instead by verifying \emph{evidence} that substantiates the correctness of the count. Evidence records and corresponds to the individual steps taken in the computation of winners 
%(That is, our approach is an instance of verifiable computation \footnote{DP: reference}). 
%The evidence directly corresponds to the individual steps of the count
(and could so also be verified, or spot-checked manually), giving transparency. We minimize the computational trust base by verifying our executables down to the machine level using a combination of HOL \cite{hol} and CakeML \cite{cake}.

@@

To begin 
%with addressing the problem, 
assume $\mathcal{P}$ is an implementation of an algorithm $\mathcal{A}$. We reasonably demand each execution of $\mathcal{P}$ on a given input $x$ consisting of ballots recorded in the election to output the winners $y$ and evidence $\omega$ as claim  for correctness of the execution.  
 Having evidence available for each execution of $\mathcal{P}$ facilitates checking correctness of the computation carried out \emph{independently} of (the source code of) $\mathcal{P}$. Therefore the original problem boils down to how one can verify the evidence of any \emph{instance of computation} with any algorithm $\mathcal{A}$.

The evidence as such must be enough informative to provide transparency of tallying and also allow  voters, or at least a large pool of scrutineers, to verify for themselves that the tally is authentically processed as per instructions of the counting scheme  so that winners truly reflect will of the voters. Therefore the question becomes what information is enough  for establishing transparency and verifiability of tallying  and what data structure for recording the information in the evidence we should choose.

@@

 \item\textbf{elected candidates} who are the already elected ones,% whose surplus votes must be transferred,

@@

In a secondary examination of the STV family, we also identify a general algorithmic mechanism for performing computation on evidence, which realises the \emph{commonalities} among the family members. This generic method comprises seven general actions%~(Figure~\ref{GenSteps})  
~(Definition~\ref{transitions}) corresponding to actions that tally officers take when counting votes.
\begin{definition}[Generic Counting Actions]\label{transitions}
 The following comprise the generic STV counting actions.
 %The generically appearing counting actions in the STV family are the followings. 
%\begin{figure}  
 \begin{itemize}
\item\textbf{count} for counting the uncounted ballots,
\item\textbf{elect} for electing all or some of the electable candidates, 
\item\textbf{transfer-elected} for distributing surplus votes of elected, %candidates, 
\item\textbf{eliminate} to exclude one or some candidates, 
\item\textbf{transfer-removed} to distribute votes of   the eliminated, %candidates, 
\item\textbf{elected-win} for terminating tallying whenever the vacancies are filled or there is no continuing candidate left, 
\item\textbf{hopeful-win} for terminating tallying whenever  the number of already elected and continuing candidates collectively does not exceed the initial vacancies. 
\end{itemize}
%\caption{Generic STV Counting Steps}
%\label{GenSteps}
%\end{figure}
\end{definition}
Each action is constituted of  preconditions on when and how the action comes into effect, and postconditions dictating what effect applying the action brings about by describing the new state to which computation proceeds. In fact, these conditions are the formal counterparts of the legal clauses in the textual specification of the counting scheme which tell tally officers how to proceed with tallying, and they are common between all flavours of STV.

For example a constant precondition for applying elimination  asserts that there must be no uncounted ballot left and another one declares that no candidate should be electable at the current state of counting. A postcondition of elimination is that once a candidate is excluded from counting, the list of continuing candidates is updated accordingly so that the excluded candidate no longer received votes. Conjunctions of such declarations comprise what an action is and what an application of that action means.

@@

\begin{definition}[The generic STV machine]\label{STVMachine}
Let $\mathcal{S}$  and $\mathcal{T}$
be respectively the set of machine states and transition labels.  Also assume for $t\in\mathcal{T}$, $S_{t}$ $=$ $\bigwedge_{\atop{i\leq j_{t}}} \psi_{i}$ where  for $i\in\{1...j_{t}\}$, $\psi_{i}$ is the formal specification (in higher-order logic) of a pre- or post-condition of $t$. Then the \emph{generic STV} is the triple $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$.  
\end{definition}
There are nonetheless variations in STV algorithms  which we  recognise  by separate \emph{instantiations} into the machine.  
\subsection{Instantiations of the Machine with Specific STV schemes}
Differences exist between counting algorithms of STV schemes.  For example in the STV used for electing the upper house representatives of the Victoria state of Australia (Victoria STV)~\cite{vic}, distributing votes of an eliminated candidate occurs step by step in order of the magnitude of the fractional values that those votes carry.  Consequently, a pre-condition for applying the action is to first assure that upon each instance of applying transfer-removed action (a) exactly those votes which have the same fractional value are transferred and (b) the transfer happens according to the magnitude order. In contrast, for example, the STV used in the lower house elections of Tasmania state of Australia (Tasmania STV) does not have such precondition simply because votes of an eliminated candidate are transferred in a single application of transfer-removed. There are also variations in postconditions. For example transfer-elected of the ACT STV~\cite{act} used in the lower house elections of the ACT state of Australia requires distributing only the last parcel of surplus votes received by an elected candidate. But Victoria STV specifies transferring all surplus votes of an elected candidate.

We formally realise variations of STV algorithms by instantiation~(Definition~\ref{STVInst}) of the generic machine. Recall that the semantics of the machine~(Definition~\ref{STVMachine}) is formed by conjunctions of formally specified universally appearing pre- and postconditions in STV algorithms. An instantiation of the machine with an STV algorithm $\mathcal{A}$ happens by enriching the generic semantics by adding formal specifications of the legal clauses that are specific to $\mathcal{A}$. We hence come to define an operational semantics functioning in accordance with instructions of $\mathcal{A}$. 
\begin{definition}[Instantiation of the Machine]\label{STVInst}
Assume the generic machine $M = \langle \mathcal{S}, \mathcal{T}, (S_t)_{t \in \mathcal{T}} \rangle$ as in Definition~\ref{STVMachine}. An instantiation $\hat{\mathcal{A}}$ for an STV algorithm $\mathcal{A}$ of  the generic STV machine $M$ is a triple $\langle \mathcal{S}, \mathcal{T}, (S'_t)_{t \in \mathcal{T}} \rangle$, where for each $t\in\mathcal{T}$, $S'_{t}$ $=$ $S_{t}\wedge\bigwedge_{\atop{i\leq j'_{t}}} \phi_{i}$. Each $\phi_{i}$ is the formal specification of a pre- or postcondition specific to $\mathcal{A}$.   
\end{definition}

The operational semantics of each instantiation allows  to perform operations such as validation on evidence. Definition~\ref{verifier} lays down what verifying an evidence amounts to. Informally speaking, to check if  evidence is valid one simply needs to inspect if transitions  between each two consecutive states of computation appearing in the evidence occur by a legitimate application of a counting action of the scheme used.

\begin{definition}[verifier]\label{verifier}
Assume $\hat{\mathcal{A}}= \langle \mathcal{S}, \mathcal{T}, (S'_t)_{t \in \mathcal{T}} \rangle$ is an instantiation of the STV algorithm $\mathcal{A}$. A verifier for instances of computation of $\mathcal{A}$ is a function $\hat{\mathcal{V}}$ mapping from $\mathbb{Q}\times\mathbb{N}\times 2^{C}\times$ \textsf{List}$(\mathcal{S})$ to $\{0,1\}$ such that for any evidence $\hat{\omega} = (Q,s,C,\Omega)$ where $\Omega=\langle\Omega_{1},\dots,\Omega_{n}\rangle$, $\hat{\mathcal{V}}(\hat{\omega}) = 1$ if and only if the following holds:\\
$\forall i\in\{1,\dots,n-1\}.$ 
%\hspace*{0.2cm}$\forall \Omega_{i-1}$ $\Omega_{i}\in\mathcal{S}.$
%$\exists \Omega_{1}$ $\Omega_{2}\in$ \textsf{List} $(\mathcal{S}).$ $\Omega = \Omega_{1} + [\Omega_{i-1};\Omega_{i}] + \Omega_{2}$  
$\exists t\in\mathcal{T}.$  $(\Omega_{i-1},\Omega_{i})\in t$ $\wedge~\vdash S'_{t}[\Omega_{i-1},\Omega_{i}]$\\ 
where $\mathbb{Q}$ and $\mathbb{N}$ respectively represent the set of rational and natural numbers, $2^{C}$ is the set of all subsets of $C$, and \textsf{List}$(\mathcal{S})$ is the set of all possible lists of machine states. $\Omega_{i-1}$ is a pre-state and $\Omega_{i}$ is a post-state visited in the execution. The last line means that pre- and post-conditions  of applying $t$ to move from $\Omega_{i-1}$ to $\Omega_{i}$  are logically true. 
\end{definition} 
We shall next provide an example to concretely illustrate a piece of evidence and exemplify how STV schemes work. 
 \subsection{An Example of a Concrete Piece of Evidence}
\label{subsec:InstEv}
Figure~\ref{EvInst} depicts evidence of a small election whose counting algorithm is the ACT STV 
with three candidates $a$, $b$ and $c$, and initially recorded ballots containing
$b_1=([a,c],1/1)$, $b_2=([a,b,c],1/1)$, $b_3=([a,c,b],1/1)$,
$b_4=([b,a],1/1)$, $b_5=([c,b,a],1/1)$.  The first component of each ballot is a preference-ordered list of candidates, and the second is the fractional transfer value (initially set to $1/1$). The evidence consists of a \emph{header} that specifies the quota (computed according to the ACT STV instructions), the number of vacancies, and the list of competing candidates. 
The fourth line is the election result, and the remainder of the evidence are non-final states visited to compute this
outcome. Each non-final state shows uncounted ballots, tallies of candidates, piles of votes attracted, backlog of the elected, backlog of the eliminated, list of elected candidates, and list of continuing candidates.

First preferences for each candidate are computed. Ballots counted in favour of particular candidates are placed onto that candidate’s pile. In this election, $a$ is the first preference on $b_{1}$, $b_{2}$, and $b_{3}$, $b$ receives $b_{4}$, and $c$ receives $b_{5}$. Tallies are updated so that tally of $a$ becomes 3, and $b$
and $c$ each reach $1/1$. As $a$ exceeds the quota, he is elected and the fractional value of his surplus is updated according to the ACT scheme. Then $a$'s surplus is transferred to continuing candidates $b$ and $c$ and votes are counted according to next preferences shown on $a$'s surplus votes. Candidate $b$ and $c$ each attracts one vote. However  neither of them reaches to the quota. Hence the weakest one $b$ is eliminated. As the number of elected and continuing candidates is equal to 
%does not exceed 
the initial vacancies, the election terminates by declaring $a$ and $c$ winners.
\begin{small} 
\begin{figure}[b]
\begin{tabular}{c@{\hspace{2cm}}c} 
\AxiomC{\tiny 8/3} \noLine

@@

\UnaryInfC{\tiny [$b_4$,([a,b,c],1/9)]; a\{3/1\} b\{10/9\} c\{11/9\}; a\{[]\} b\{[]\}
c\{[[$b_{5}$,([c],1/9),([c,b],1/9),([c],1/9)]]\}; []; [a]; [c]}

@@

   \node[state,
  below of= STV1,
  xshift = - 1cm,
  node distance=1.2cm,
  anchor=center] (T) 
 {%
 \begin{tabular}{l}
  \textbf{Translation}
 \end{tabular}
 };

@@

We also have another motivation for developing the auxiliary module through this three-part process of specification, implementation and then verification. Familiarity, but no expertise, with our framework and its general purpose is needed for extending it to synthesise an executable verifier for one's desired STV scheme. On the one hand, an average user may lack enough skills in grasping functional programming style which we rely on for implementing verifiers in HOL4. However, the same user most probably has had exposure to formulations of mathematical properties  in first-order logic syntax. Including purely descriptive logical assertions as our specifications, proven equivalent to their implementations, facilitates the users with means to understand the functionality of components, modules and the system as a whole simply by inspecting the specifications instead of implementations.
\subsection{Data Structure of Machine States}
\label{MachineData}
 We choose the data structure  given in Definition~\ref{judgement} for implementing the abstract data type underlying the generic STV.  We have four reasons for this choice of data structure. 
\item As the abstract syntactic representation of evidence closely represents concrete evidences, designing the parser and inspecting  its  correctness is less challenging.  
 \item HOL4 has well developed libraries comprised of verified assertions of operations on list structure. By structuring the data on lists, we facilitate us and users  with exploitation of  already verified tools to formalise the framework and avoid inventing the unnecessary from the scratch. 
 \item HOL4 also has well-developed tactics for discharging proof obligations on assertions involving list structure and operations on it. We therefore come to provide us with means for ease of verification of the formalised assertions.
 \item Understanding the data structure used in the  framework implementation is critical for ease of  its usability and extensibility by third parties.  The type $\mathit{judgement}$ closely models concrete evidences such as the one in figure~\ref{EvInst}. Therefore one can sensibly perceive the  abstraction step taken for modelling  concrete data which in turn enhances understandability of the framework and its mechanism. 
\begin{definition}\label{judgement}
We formalise machine states as an inductive type  $\mathit{judgement}$ whose constructors are  $\mathit{NonFinal}$ and $\mathit{Final}$. 
 A $\mathit{Final~(w)}$ judgement declares $w$ as winners of the election. A  $\mathit{NonFinal~(ba,t,p,bl_{1},bl_{2},e,h)}$ judgement  consists of uncounted ballots $ba$, tally $t$, pile $p$, $bl_{1}$ the backlog of elected candidates, $bl_{2}$ the backlog of eliminated candidates, and $e$ and $h$ for the list of elected and continuing~(hopeful) candidates. 
\HOLTyOp{judgement}\;=\\
\;\;\;\;\HOLConst{NonFinal}\;(\HOLTyOp{ballots}\;\HOLTokenProd{}\;\HOLTyOp{tallies}\;\HOLTokenProd{}\;\HOLTyOp{piles}\;\HOLTokenProd{}\\\;\;\;\;\;\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list}\;\HOLTokenProd{}\;\HOLTyOp{cand}\;\HOLTyOp{list})\\
\;\;\HOLTokenBar{}\;\HOLConst{Final}\;(\HOLTyOp{cand}\;\HOLTyOp{list})
 The types  $\mathit{ballots}$, $\mathit{tallies}$ and $\mathit{piles}$ are respectively abbreviations for  $\mathit{((cand ~list)\times rat) list}$,   $\mathit{(cand\times rat)~list}$ and $\mathit{(cand\times(ballots~list)~list) ~list}$ where $\mathit{rat}$ is the HOL4 type of fractional numbers.
 Recall that objects whose type is $\mathit{piles}$ serve as containers for recording the ballots allocated to each candidate. Our choice for type of piles allows us to store the ballots received by each candidate in chunks of lists rather than one single list containing all of them.  We have two reasons for designing the type of piles as it stands instead of $\mathit{(cand\times(ballots~list))~list}$;  
\paragraph*{(a)} Some STV schemes such as the lower house ACT and Tasmania STV, employ a notion called ''last parcel''. In short the last parcel of a candidate is the collection of those ballots received by the candidate at the last round of application of the count transition which made the tally of the candidate reach or exceed the quota and therefore be elected.  Only the last parcel, instead of all, of an elected candidate is distributed, according to next preferences shown on the ballots in the last parcel, to the continuing candidates.
%the ballots received by the candidate.
 Also  they update the fractional transfer value, that each ballot in the last parcel  awaiting transfer carries, based on the length of the last parcel. To  accommodate this notion and its effect, we need identifying the ballots constituting the last parcel of an elected. We hence formalise the pile to record, for each  continuing candidate, a list consisting of lists of ballots each received upon  chronological applications of count transition.  Once a candidate is elected their pile may be manipulated differently from the continuing candidates. 
\paragraph*{(b)} We noted earlier that the STV used for the upper house elections in the Victoria state transfers votes of an eliminated candidate chunk by chunk in several applications of transfer-removed, rather than in a single action, in order of the magnitude of the fractional value that the chunks carry. We therefore need to rearrange the pile of ballots of an eliminated candidate into lists based on equality of the fractional value of ballots and then distribute them in the proper order as the scheme requires. This means that the type of piles has to be as it is defined above.   
Implementing piles in this way enables us to tailor the semantics of the transfer and elect transitions and their instantiations in such a way to modularly formalise several STV schemes which use the last parcel or stepwise distribution of votes form eliminated candidates or the combination of both.

@@

The predicate ValidInitCandList which realises item (6) asserts that each candidate has a pile and a tally. We also declare that elements in the first component of the tally $t$ and pile $p$ are distinct. Therefore we come to satisfy item (1).
